---
format: 
  revealjs:
    slide-number: true
    theme: [default, ../theme/presentation_theme.scss, ../theme/presentation_components.scss]
    embed-resources: true
    preview-links: false
    output-file: "presentation"
    output-ext: "html"
    highlight-style: github-light
    include-in-header: "../theme/fonts-include.html"
filters:
  - ../theme/rn-shorthand.lua
  - ../_extensions/EmilHvitfeldt/roughnotation/rough.lua
---

```{r}
#| label: setup
#| include: false
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  out.width = "100%"
)

if (!require("here")) install.packages("here")

library(here)

here::i_am("Presentation/presentation.qmd")

if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggbeeswarm")) install.packages("ggbeeswarm")
if (!require("ggview")) install.packages("ggview")
if (!require("magick")) install.packages("magick")
if (!require("cowplot")) install.packages("cowplot")
if (!require("remotes")) install.packages("remotes")
if (!require("waffle")) remotes::install_github("hrbrmstr/waffle")

library(tidyverse)
library(ggbeeswarm)
library(ggview)
library(magick)
library(cowplot)
library(waffle)

path_materials <-
  "Presentation/Materials"

include_local_figure <- function(data_source) {
  knitr::include_graphics(
    path = here::here(
      path_materials,
      data_source
    ),
    error = TRUE
  )
}
```

```{r}
#| label: dynamic link generation
#| include: false
project_name <-
  basename(here::here())

display_name <-
  paste("Biostatistika", "-", project_name)

github_url <-
  paste0(
    "https://github.com/CUNI-NATUR-Biostatistics/",
    project_name
  )

slides_url <-
  paste0(
    "https://CUNI-NATUR-Biostatistics.github.io/",
    project_name, "/"
  )
```

## Co nám může říct měření tučňáků? {.course-title .bg-indigo}

::: {.course-title-mark}
![Logo kurzu Biostatistika](../theme/course-logo-reversed.svg) BIOSTATISTIKA
:::

::: {.course-title-number}
LXX
:::

::: {.course-title-formal}
Vztahy dvou proměnných
:::

::: {.course-title-link}
[cuni-natur-biostatistics.github.io/materialy.html](https://cuni-natur-biostatistics.github.io/materialy.html)
:::

::: {.notes}
V každé nové lekci nahraďte ukázkovou otázku, název a označení lekce vlastními údaji. Otázka na titulním snímku musí vycházet ze schválené datové linky lekce.
:::

## Co vidíme v měření tučňáků?

TODO: Ukažte jeden konkrétní obrázek, měření nebo jiný háček k úvodní otázce. Nechte studenty nejprve popsat, co vidí.

::: {.notes}
Tento ukázkový snímek nahraďte prvním datovým nebo biologickým háčkem vlastní lekce; ponechte vazbu na otázku z titulního snímku.
:::

## Výsledky učení

Po absolvování této přednášky budete schopni:

- TODO: výsledek učení 1
- TODO: výsledek učení 2
- TODO: výsledek učení 3

# Obsah přednášky

## Snímek 1

TODO: obsah snímku

```{r}
#| label: demo-plot
#| echo: true
# TODO: ukázka kódu v R
```

## Shrnutí

- TODO: klíčové body přednášky

# Slide Creator's Guide {.center}

[Slides in this section are intended for lecture authors, not students. Delete this section and the `# Tools Overview` section before using as an actual lecture.]{.text-size-smaller .text-color-grey_olive}

# Golden Rules {.center}

## Rule 1 — Data story first {.bg-parchment}

::: columns
::: {.column width="48%"}
::: {.card-warning}
[Wrong]{.card-title}
**Pearson correlation**

Pearson's correlation coefficient measures the strength of the linear relationship between two variables...

[formula]
:::
:::
::: {.column width="48%"}
::: {.card-insight}
[Right]{.card-title}
**How do we turn the shape of a cloud into a number?**

[scatter plot the students have seen for several slides already]

*What properties would such a number need to have?*

→ **Correlation coefficient** [the name arrives as the answer]{.text-size-smaller .text-color-grey_olive}
:::
:::
:::

[A statistical term never opens a concept as a heading. Data first → observation → question → only then the name.]{.text-size-smaller .text-color-grey_olive}

## Rule 2 — One idea per slide

**Every slide communicates one thing.**

If you need to say two things, use two slides.

If you need to say three things, add an interaction between them.

::: {.takeaway-strip}
**One message. One slide. Contribute to understanding — not to overload.**
:::

[Test: cover the heading — is the main idea of the slide still clear from the visual or a single sentence?]{.text-size-smaller .text-color-grey_olive}

## Rule 3 — Interaction every 10–15 minutes

::: {.panel-note}
[Six interaction types]{.panel-title}

| Type | When to use |
|---|---|
| Prediction question | Before revealing a result, graph, or test output |
| Conceptual MCQ | To surface a common misconception; ideally with a polling tool |
| Think–Pair–Share | After a new concept; 30 s thinking + 1–2 min pair discussion |
| Silent write | Before discussion or reveal; each student writes one sentence |
| Interpretation task | After showing a graph, model output, or table |
| Data exploration | To open a topic — what patterns / hypotheses do you see? |
:::

::: {.panel-question}
[Example — prediction question]{.panel-title}

*How does the confidence interval change if we double the sample size?*

**A)** Narrows   **B)** Stays the same   **C)** Widens
:::

## Rule 4 — Always frame biologically

Every statistical method must be grounded in a concrete biological question.

::: {.blockquote}
**Four steps of the biological frame:**

1. **Biological question** — "we want to find out whether…"
2. **Data** — description, type, source
3. **Statistical method** — what and why
4. **Interpretation in biological context** — what the result says about the world, not about numbers
:::

[Never start with "today we'll learn the t-test". Always start with "we want to know whether…"]{.text-size-smaller .text-color-grey_olive}

## Rule 5 — Alternate visual patterns (3-of-5 rule)

**At most 3 of 5 consecutive slides may use a named component as the primary visual element.**

A plain slide — a bold question on an empty background, a full-width graph, a single prominent number — carries more weight precisely because the surrounding slides are more structured.

::: {.strip-orange}
**3 of 5 — the law of slide visual contrast** · review the block before finalising
:::

[Components: `.card-*`, `.panel-*`, `.blockquote`, `.takeaway-strip`. If you reach 4+ in a row, convert one slide to a plain visual or an unframed question.]{.text-size-smaller .text-color-grey_olive}

# Five Slide Families {.center}

## Five slide families

::: columns
::: {.column width="62%"}

| Family | Goal | Typical component |
|---|---|---|
| **Interaction** | Prediction / MCQ / TPS | bold question, `.panel-question` |
| **Data reveal** | New data or output | full-width graph, plain slide |
| **Interpretation** | Read the result | `.panel-result`, `.takeaway-strip` |
| **Rule / warning** | Heuristic, error | `.card-warning`, `.panel-warning` |
| **Transition / summary** | Reset, conclusion | `.blockquote` in two columns |

:::
::: {.column width="35%"}
::: {.card-olive}
[Variety of families]{.card-title}
In one lecture, cycle through all five families.

No block of 5+ slides should belong to only one family.
:::
:::
:::

[Students should be able to recognise the role of a slide before reading the text — from the colour key, layout, or position in the structure.]{.text-size-smaller .text-color-grey_olive}

# Model output for authors {.center}

## Build model tables from the fit

Use this pattern only after a lesson has introduced the model and its base-R output. Keep the model fit and preparation in hidden chunks near the first table; show only the selected Czech-labelled table. Adjust terms, quantities, rounding, and units to the actual teaching question.

```r
# Hidden authoring code after fitting mod_priklad locally:
koeficienty <-
  broom::tidy(mod_priklad, conf.int = TRUE)

tabulka <-
  data.frame(
    Parametr = koeficienty$term,
    Odhad = koeficienty$estimate,
    `Standardní chyba` = koeficienty$std.error,
    check.names = FALSE
  )

tinytable::tt(tabulka, digits = 2)
```

Do not copy coefficient values into a hand-written table. Compare the generated values with `coef(mod_priklad)` and `summary(mod_priklad)` and inspect the rendered HTML and PDF before reusing this example.

# Tools Overview {.center}

## `.box-*` aliases — when to use what

| Alias | Maps to | Weight | Colour | When to use |
|-------|-----------|------|-------|------------|
| `.box-question` | `.panel-question` | panel (lighter) | orange left border | MCQ, reflection question |
| `.box-result` | `.panel-result` | panel (lighter) | indigo left border | main conclusion, analysis result |
| `.box-data` | `.panel-data` | panel (lighter) | olive left border | dataset description, data context |
| `.box-rule` | `.card-rule` | card (heavier) | amethyst top strip + shadow | rule, heuristic |
| `.box-warning` | `.card-warning` | card (heavier) | red top strip + shadow | methodological warning |

[Panel aliases (lighter): coloured left border, no shadow. Card aliases (heavier): coloured top strip + shadow. Inner heading: `[text]{.panel-title}` for panels · `[text]{.card-title}` for cards.]{.text-size-smaller .text-color-grey_olive}

## `.box-question` — interaction, MCQ

::: columns
::: {.column width="55%"}
::: {.box-question}
[Question for reflection]{.panel-title}

What percentage of penguin body-mass variability is explained by bill length?

**A)** less than 30 % &nbsp;&nbsp; **B)** 30–60 % &nbsp;&nbsp; **C)** more than 60 %
:::
:::
::: {.column width="42%"}
```markdown
::: {.box-question}
[Question for reflection]{.panel-title}

Question text + choices A / B / C.
:::
```

[= `.panel-question` · orange left border · heading `{.panel-title}`]{.text-size-smaller .text-color-grey_olive}
:::
:::

## `.box-result` — main result

::: columns
::: {.column width="55%"}
::: {.box-result}
[What we see]{.panel-title}

Each millimetre of bill length corresponds on average to **+8.7 g** body mass
(95% CI: 6.2–11.2 g). The model explains 45 % of body-mass variability.
:::
:::
::: {.column width="42%"}
```markdown
::: {.box-result}
[What we see]{.panel-title}

Main conclusion or analysis result.
:::
```

[= `.panel-result` · indigo left border · heading `{.panel-title}` or `{.panel-lead}`]{.text-size-smaller .text-color-grey_olive}
:::
:::

## `.box-data` — dataset description

::: columns
::: {.column width="55%"}
::: {.box-data}
[Palmer Penguins]{.panel-title}

344 penguins from three islands · 3 species *(Adelie, Chinstrap, Gentoo)*

Variables: bill length and depth (mm), flipper length (mm), body mass (g), sex, year
:::
:::
::: {.column width="42%"}
```markdown
::: {.box-data}
[Dataset name]{.panel-title}

Brief description: number of
observations, variables, source.
:::
```

[= `.panel-data` · olive left border · heading `{.panel-title}`]{.text-size-smaller .text-color-grey_olive}
:::
:::

## `.box-warning` — methodological warning

::: columns
::: {.column width="55%"}
::: {.box-warning}
[Caution — correlation ≠ causation]{.card-title}

A longer bill does not imply that *the bill causes* greater body mass — both traits
may reflect overall body size.
:::
:::
::: {.column width="42%"}
```markdown
::: {.box-warning}
[Warning heading]{.card-title}

Methodological error, misinterpretation,
data limitation.
:::
```

[= `.card-warning` · **card** · red top strip + shadow · heading `{.card-title}`]{.text-size-smaller .text-color-grey_olive}
:::
:::

## `.box-rule` — rule, heuristic

::: columns
::: {.column width="55%"}
::: {.box-rule}
[Mean or median?]{.card-title}

If the data are symmetric → mean.

If the data are skewed or contain outliers → median.
:::
:::
::: {.column width="42%"}
```markdown
::: {.box-rule}
[Rule heading]{.card-title}

Condition → recommendation.
Or a verbal heuristic.
:::
```

[= `.card-rule` · **card** · amethyst top strip + shadow · heading `{.card-title}`]{.text-size-smaller .text-color-grey_olive}
:::
:::

## Combination: `.box-question` + `.takeaway-strip`

::: columns
::: {.column width="55%"}
::: {.box-question}
[Question for reflection]{.panel-title}

Why does bill length differ between penguin species and within a single species?
:::

::: {.takeaway-strip}
**Morphology as adaptation.** Bill length reflects life strategy — Gentoo penguins forage differently from Adelie.
:::
:::
::: {.column width="42%"}
```markdown
::: {.box-question}
[Question]{.panel-title}
Question...
:::

::: {.takeaway-strip}
**Conclusion.** Closing the conclusion.
:::
```

[`.takeaway-strip` after `.box-question` closes a slide sequence. One slide = one question + one conclusion.]{.text-size-smaller .text-color-grey_olive}
:::
:::

## Fragment: incremental reveal of `.box-result`

::: columns
::: {.column width="55%"}
Let's look at the linear model output.

::: {.fragment}
::: {.box-result}
[Result]{.panel-title}

Bill length coefficient: **+8.7 g/mm** (95% CI: 6.2–11.2 g).
The effect is statistically and biologically significant.
:::
:::
:::
::: {.column width="42%"}
```markdown
Sentence before the pause.

::: {.fragment}
::: {.box-result}
[Result]{.panel-title}
Text that is revealed.
:::
:::
```

[`::: {.fragment}` wraps any `.box-*` block. Students see the sentence → click → result.]{.text-size-smaller .text-color-grey_olive}
:::
:::

## Tool: plain question without a frame {.center}

<br>

::: {.text-center .text-size-large .text-bold}
How does the confidence interval change if we double the sample size?
:::

<br>

::: {.text-center}
**A)** Narrows &nbsp;&nbsp;&nbsp; **B)** Stays the same &nbsp;&nbsp;&nbsp; **C)** Widens
:::

<br>

[Plain slide: bold question + choices without a frame. Maximum visual contrast against slides with `.box-*` components.]{.text-size-smaller .text-color-grey_olive}

## Roughnotation — use sparingly

Each millimetre of penguin bill length corresponds on average to
[+8.7 g body mass]{.rn-underline-indigo} — the effect is
[biologically significant]{.rn-box-orange}.

<br>

::: columns
::: {.column width="48%"}
```markdown
[text]{.rn-underline-indigo}
[text]{.rn-box-orange}
[text]{.rn-circle-amethyst}
[text]{.rn-bracket-indigo}
```
:::
::: {.column width="48%"}
```markdown
[text]{.rn-highlight-orange}
[text]{.rn-cross-graphite}
```

Colours: `indigo`, `orange`, `amethyst`,
`olive`, `graphite`, `red`
:::
:::

[**Sparingly:** max 1–2 effects per slide, 3–5 per lecture. Types: `underline`, `box`, `circle`, `bracket`, `highlight`, `cross`. Not as decoration — only for one key moment in a concept.]{.text-size-smaller .text-color-grey_olive}

# Reference Slides: SCSS Components {.center}

# Cards {.center}

## Card: basic structure

::: {.card-indigo}
[Card heading]{.card-title}

Card body — one or two short paragraphs. The font is intentionally smaller than the base
slide size (0.82 em) to fit more content.

- Cards can also contain a list
- Each item on its own line
:::

[Syntax: `::: {.card-indigo}` + `[text]{.card-title}` for the coloured top strip — without the old `.slide-card` + variant combination.]{.text-size-smaller .text-color-grey_olive}

## Card: .card-title

::: columns
::: {.column width="60%"}
::: {.card-insight}
[This is .card-title]{.card-title}

`[text]{.card-title}` must be the **first thing** inside the card — it creates the coloured top strip.
:::
:::

::: {.column width="38%"}
```markdown
::: {.card-insight}
[Section name]{.card-title}

Card body.
:::
```
:::
:::

## Card: compact variant

::: columns
::: {.column width="48%"}
::: {.card-orange}
[Standard]{.card-title}

`.card-orange` — padding `0.7rem 1rem`. Suitable for longer content.
:::
:::

::: {.column width="48%"}
::: {.card-orange .card-compact}
[Compact]{.card-title}

`.card-orange .card-compact` — reduced padding (`0.45rem 0.75rem`). For denser layouts.
:::
:::
:::

## Card: overview of variants

::: columns
::: {.column width="18%"}
::: {.card-indigo}
[indigo]{.card-title}
`.card-indigo`
:::
:::
::: {.column width="18%"}
::: {.card-orange}
[orange]{.card-title}
`.card-orange`
:::
:::
::: {.column width="18%"}
::: {.card-amethyst}
[amethyst]{.card-title}
`.card-amethyst`
:::
:::
::: {.column width="18%"}
::: {.card-olive}
[olive]{.card-title}
`.card-olive`
:::
:::
::: {.column width="18%"}
::: {.card-red}
[red]{.card-title}
`.card-red`
:::
:::
:::

::: columns
::: {.column width="16%"}
::: {.card-question}
[question]{.card-title}
`.card-question`
:::
:::
::: {.column width="16%"}
::: {.card-insight}
[insight]{.card-title}
`.card-insight`
:::
:::
::: {.column width="16%"}
::: {.card-warning}
[warning]{.card-title}
`.card-warning`
:::
:::
::: {.column width="16%"}
::: {.card-data}
[data]{.card-title}
`.card-data`
:::
:::
::: {.column width="16%"}
::: {.card-result}
[result]{.card-title}
`.card-result`
:::
:::
::: {.column width="16%"}
::: {.card-rule}
[rule]{.card-title}
`.card-rule`
:::
:::
:::

[Semantic variants are aliases: `.card-question` = orange, `.card-insight` = `.card-result` = indigo, `.card-warning` = red, `.card-data` = olive, `.card-rule` = amethyst. Shorthand: `.box-question`, `.box-result`, `.box-data`, `.box-warning`, `.box-rule`.]{.text-size-smaller .text-color-grey_olive}

# Panels {.center}

## Panel: `.panel-result`

::: columns
::: {.column width="55%"}
::: {.panel-result}
[Key insight]{.panel-title}

Purple left border. For the main message of a block — conclusion, analysis result, or key number.
:::
:::
::: {.column width="42%"}
```markdown
::: {.panel-result}
[Heading]{.panel-title}
Panel text.
:::
```

[`.panel-title` is `display: block`; gives the heading h3 appearance without `###`.]{.text-size-smaller .text-color-grey_olive}
:::
:::

## Panel: `.panel-question`

::: columns
::: {.column width="55%"}
::: {.panel-question}
[Conclusion or option]{.panel-title}

Orange left border. For conclusions, comparison results, or offering multiple answer choices.
:::
:::
::: {.column width="42%"}
```markdown
::: {.panel-question}
[Conclusion or option]{.panel-title}
Panel text.
:::
```
:::
:::

## Panel: `.panel-question` + `.panel-lead`

::: columns
::: {.column width="55%"}
::: {.panel-question}
[Computation result]{.panel-lead}

Large highlighted result — replaces the old `.answer-title`. Use at most once per slide.
:::
:::
::: {.column width="42%"}
```markdown
::: {.panel-question}
[Computation result]{.panel-lead}

Text below the result.
:::
```

[`.panel-lead` is large and bold. Suitable for a key number or conclusion.]{.text-size-smaller .text-color-grey_olive}
:::
:::

## Panel: `.panel-note` and `.panel-warning`

::: columns
::: {.column width="48%"}
::: {.panel-note}
[Instructions or output]{.panel-title}

Neutral olive border. For instructions, workflow steps, or model output.
:::
:::
::: {.column width="48%"}
::: {.panel-warning}
[Caution!]{.panel-title}

Red left border. For warnings, common errors, or deliberate alerts.
:::
:::
:::

## Panel: overview of colour variants

::: columns
::: {.column width="18%"}
::: {.panel-indigo}
[indigo]{.panel-title}
`.panel-indigo`
:::
:::
::: {.column width="18%"}
::: {.panel-orange}
[orange]{.panel-title}
`.panel-orange`
:::
:::
::: {.column width="18%"}
::: {.panel-amethyst}
[amethyst]{.panel-title}
`.panel-amethyst`
:::
:::
::: {.column width="18%"}
::: {.panel-olive}
[olive]{.panel-title}
`.panel-olive`
:::
:::
::: {.column width="18%"}
::: {.panel-red}
[red]{.panel-title}
`.panel-red`
:::
:::
:::

[Semantic variants are aliases: `.panel-result` = indigo, `.panel-question` = orange, `.panel-warning` = red, `.panel-note` = amethyst, `.panel-data` = olive. Shorthand (`.box-*`) wraps the matching panel.]{.text-size-smaller .text-color-grey_olive}

# Takeaway strip {.center}

## Takeaway strip: standard

Sentence that leads into the conclusion — contextual paragraph above the strip.

::: {.takeaway-strip}
**Main slide conclusion.** `.takeaway-strip` — indigo-to-orange gradient background,
indigo left border. One per slide, placed at the bottom.
:::

[`::: {.takeaway-strip}` — the gradient is built in, no extra class needed.]{.text-size-smaller .text-color-grey_olive}

## Takeaway strip: compact variant

Sentence above the strip as usual.

::: {.takeaway-strip}
Main conclusion — standard strip.
:::

::: {.takeaway-strip .strip-compact}
Supplementary conclusion — `.strip-compact` reduces `margin-top`. Useful for two conclusions together.
:::

## Strip: colour variants

::: columns
::: {.column width="18%"}
::: {.strip-indigo}
`.strip-indigo`
:::
:::
::: {.column width="18%"}
::: {.strip-orange}
`.strip-orange`
:::
:::
::: {.column width="18%"}
::: {.strip-amethyst}
`.strip-amethyst`
:::
:::
::: {.column width="18%"}
::: {.strip-olive}
`.strip-olive`
:::
:::
::: {.column width="18%"}
::: {.strip-red}
`.strip-red`
:::
:::
:::

# Blockquote {.center}

## Blockquote div: `.blockquote`

::: {.blockquote}
**Definition or quotation.** `.blockquote` — light purple background, indigo left border (6px),
rounded corners. Font 0.82 em.

- Can contain a list
- Each item on its own line
:::

[Use `::: {.blockquote}`, not Markdown `>`. The div variant gets correct styling;
Markdown blockquote gets different styles from `presentation_theme.scss`.]{.text-size-smaller .text-color-grey_olive}

# Slide backgrounds {.center}

## Backgrounds: dark variants {.bg-indigo}

This slide has the class `{.bg-indigo}`. Text and headings are automatically white.

::: columns
::: {.column width="48%"}
Syntax — add the class to the slide heading:

```markdown
## Slide title {.bg-indigo}
```
:::
::: {.column width="48%"}
Available dark variants:

- `{.bg-indigo}` — primary indigo
- `{.bg-amethyst}` — lighter purple
- `{.bg-olive}` — grey-green
- `{.bg-red}` — red
- `{.bg-graphite}` — dark grey
:::
:::

## Backgrounds: light variants {.bg-parchment}

This slide has the class `{.bg-parchment}`. Text is automatically graphite.

::: columns
::: {.column width="48%"}
Available light variants:

- `{.bg-orange}` — orange
- `{.bg-parchment}` — cream
- `{.bg-white}` — pure white
:::
::: {.column width="48%"}
Text and heading colour adapts automatically — no need to add `.text-color-*`.
:::
:::

# Inline highlights {.center}

## Inline highlight: `.text-hgl-{color}`

The model intercept is −153.4 and the length coefficient is
[4.05 g/mm]{.text-hgl-indigo_velvet} — each millimetre of length corresponds on average
to four grams of body mass.

Caution — this procedure works [only when the normality assumption is met]{.text-hgl-orange}.

[Syntax: `[text]{.text-hgl-indigo_velvet}` — coloured chip with contrasting text. Available colours: `indigo_velvet`, `orange`, `amethyst`, `grey_olive`, `graphite`, `white`, `light_gray`, `parchment`.]{.text-size-smaller .text-color-grey_olive}

## Inline highlight: `.text-bg-{color}` and `.text-color-{color}`

::: columns
::: {.column width="48%"}
[`.text-hgl-indigo_velvet`]{.text-hgl-indigo_velvet} — chip with contrasting text

[`.text-bg-indigo_velvet`]{.text-bg-indigo_velvet} — chip without text colour change

[`.text-color-indigo_velvet`]{.text-color-indigo_velvet} — text colour only

[`.text-color-indigo_velvet .text-bold`]{.text-color-indigo_velvet .text-bold} — colour + bold
:::
::: {.column width="48%"}
```markdown
[text]{.text-hgl-indigo_velvet}
[text]{.text-bg-indigo_velvet}
[text]{.text-color-indigo_velvet}
[text]{.text-color-indigo_velvet .text-bold}
```
:::
:::

# Roughnotation {.center}

## Roughnotation: shorthand syntax

Shorthand `{.rn-{type}-{colour}}` expands via `rn-shorthand.lua`:

::: columns
::: {.column width="48%"}
**Short form (new):**

Each millimetre of length corresponds to approximately
[four grams of body mass]{.rn-underline-indigo}.

```markdown
[text]{.rn-underline-indigo}
```
:::
::: {.column width="48%"}
**Long form (still functional):**

```markdown
[text]{.rn-fragment rn-type=underline
  rn-color="#5D2890"}
```
:::
:::

[Available types: `underline`, `box`, `circle`, `highlight`, `cross`, `bracket`. Colours: `indigo`, `orange`, `amethyst`, `olive`, `graphite`, `red`.]{.text-size-smaller .text-color-grey_olive}

## Roughnotation: `box`, `circle`, `bracket`

::: columns
::: {.column width="32%"}
**box**

**R² = [0.76]{.rn-box-indigo}**

```markdown
[0.76]{.rn-box-indigo}
```
:::
::: {.column width="32%"}
**circle**

The relationship is [strong]{.rn-circle-orange}.

```markdown
[strong]{.rn-circle-orange}
```
:::
::: {.column width="32%"}
**bracket**

**ŷ = β₀ + [β₁x]{.rn-bracket-indigo}**

```markdown
[β₁x]{.rn-bracket-indigo}
```
:::
:::

[Animation fires on fragment arrival. Do not use both colours on the same slide.]{.text-size-smaller .text-color-grey_olive}

# Fragments and progressive disclosure {.center}

## Fragment: `.fragment` block

A sentence that is on the slide from the start.

::: {.fragment}
::: {.panel-result}
[Conclusion arrives after pause]{.panel-title}

This entire block appears as a single fragment — the student sees the sentence above
and only after clicking does the panel with the conclusion reveal itself.
:::
:::

[Syntax: `::: {.fragment}` wraps any block. Works with panels, blockquotes, strips, and cards.]{.text-size-smaller .text-color-grey_olive}

## Fragment: `.incremental` list

Which properties must a good correlation coefficient satisfy?

::: {.incremental}
- It must be symmetric: correlation of X with Y = correlation of Y with X.
- It must be dimensionless: independent of the scale of variables.
- It must be in the interval −1 to +1.
- +1 = perfect positive dependence, −1 = perfect negative dependence.
:::

[Each item appears on click.]{.text-size-smaller .text-color-grey_olive}

# Typography {.center}

## Typography: sizes

::: columns
::: {.column width="55%"}
[`.text-size-heading` — 140 % of base]{.text-size-heading}

[`.text-size-large` — 130 % of base]{.text-size-large}

Normal text (100 %, base 30 px)

[`.text-size-smaller` — 82 % of base]{.text-size-smaller}

[`.text-size-tiny` — 65 % of base]{.text-size-tiny}
:::
::: {.column width="42%"}
```markdown
[text]{.text-size-heading}
[text]{.text-size-large}
[text]{.text-size-smaller}
[text]{.text-size-tiny}
```

[`.text-size-body` = 100 % (explicit reset). Old `.text-larger`/`.text-smaller`/`.text-tiny` are removed.]{.text-size-smaller .text-color-grey_olive}
:::
:::

## Typography: weight, alignment, font

::: {.text-center}
`.text-center` — horizontal centre
:::

::: {.text-right}
`.text-right` — right alignment
:::

<br>

[`.text-font-heading` — Source Sans 3, heading weight]{.text-font-heading}

[`.text-font-monospace` — JetBrains Mono]{.text-font-monospace}

[`.text-bold` — bold]{.text-bold}

# Colour utilities {.center}

## Colours: indigo_velvet and amethyst

::: columns
::: {.column width="48%"}
::: {.bg-indigo_velvet style="padding:0.55rem 0.8rem;border-radius:10px;margin-bottom:0.5rem;"}
[`.bg-indigo_velvet` — primary indigo]{.text-color-white}
:::
[`.text-color-indigo_velvet`]{.text-color-indigo_velvet}

[`.text-hgl-indigo_velvet`]{.text-hgl-indigo_velvet}
:::
::: {.column width="48%"}
::: {.bg-amethyst style="padding:0.55rem 0.8rem;border-radius:10px;margin-bottom:0.5rem;"}
[`.bg-amethyst` — lighter purple]{.text-color-white}
:::
[`.text-color-amethyst`]{.text-color-amethyst}

[`.text-hgl-amethyst`]{.text-hgl-amethyst}
:::
:::

## Colours: orange and graphite

::: columns
::: {.column width="48%"}
::: {.bg-orange style="padding:0.55rem 0.8rem;border-radius:10px;margin-bottom:0.5rem;"}
[`.bg-orange` — vivid orange]{.text-color-graphite}
:::
[`.text-color-orange`]{.text-color-orange}

[`.text-hgl-orange`]{.text-hgl-orange}
:::
::: {.column width="48%"}
::: {.bg-graphite style="padding:0.55rem 0.8rem;border-radius:10px;margin-bottom:0.5rem;"}
[`.bg-graphite` — dark base]{.text-color-white}
:::
[`.text-color-graphite`]{.text-color-graphite}

[`.text-hgl-graphite`]{.text-hgl-graphite}
:::
:::

## Colours: neutral palette tones

::: columns
::: {.column width="23%"}
::: {.bg-parchment style="padding:0.5rem 0.7rem;border-radius:10px;border:1px solid #ccc;"}
[`.bg-parchment`]{.text-color-graphite .text-size-smaller}
:::
[`.text-color-parchment`]{.text-color-parchment .text-size-smaller .bg-graphite style="padding:1px 4px;border-radius:3px;"}

[`.text-hgl-parchment`]{.text-hgl-parchment .text-size-smaller}
:::
::: {.column width="23%"}
::: {.bg-grey_olive style="padding:0.5rem 0.7rem;border-radius:10px;"}
[`.bg-grey_olive`]{.text-color-white .text-size-smaller}
:::
[`.text-color-grey_olive`]{.text-color-grey_olive .text-size-smaller}

[`.text-hgl-grey_olive`]{.text-hgl-grey_olive .text-size-smaller}
:::
::: {.column width="23%"}
::: {.bg-light_gray style="padding:0.5rem 0.7rem;border-radius:10px;border:1px solid #ccc;"}
[`.bg-light_gray`]{.text-color-graphite .text-size-smaller}
:::
[`.text-color-light_gray`]{.text-color-light_gray .text-size-smaller .bg-graphite style="padding:1px 4px;border-radius:3px;"}

[`.text-hgl-light_gray`]{.text-hgl-light_gray .text-size-smaller}
:::
::: {.column width="23%"}
::: {.bg-white style="padding:0.5rem 0.7rem;border-radius:10px;border:1px solid #ccc;"}
[`.bg-white`]{.text-color-graphite .text-size-smaller}
:::
[`.text-color-white`]{.text-color-white .text-size-smaller .bg-graphite style="padding:1px 4px;border-radius:3px;"}

[`.text-hgl-white`]{.text-hgl-white .text-size-smaller}
:::
:::

# Layout utilities {.center}

## Layout: `.columns.align-top`

::: {.columns .align-top}
::: {.column width="48%"}
**Left column** starts at the top — without `.align-top` the content would be vertically centred.

::: {.card-indigo}
Card in the left column.

Second paragraph inside the card — the left column is taller.
:::
:::
::: {.column width="48%"}
**Right column** is shorter.

Without `.align-top` the right text would shift down to the centre of the column pair.
:::
:::

[Syntax: `::: {.columns .align-top}` — add the class `.align-top` to `.columns`.]{.text-size-smaller .text-color-grey_olive}

## Layout: vertical offsets

::: {.card-question}
[Example: `.slide-margin-top-15`]{.kicker}

This card would have `margin-top: 15%` if it had the class `.slide-margin-top-15`.
Useful for vertically centring an isolated element on a slide.
:::

::: {.takeaway-strip .strip-compact}
`.slide-margin-top-15` → `margin-top: 15%` · `.slide-margin-top-25` → `margin-top: 25%` · `.center-vertical` → absolute centre
:::

[For a group of related elements, prefer the normal-flow `.layout-stack.layout-center-y` utility shown below.]{.text-size-smaller .text-color-grey_olive}

## Layout: `.layout-grid-2`

:::: {.layout-grid-2}
::: {.panel-result}
[Left result]{.panel-title}

Two equal-width elements form one comparison.
:::

::: {.panel-question}
[Right question]{.panel-title}

The shared gap keeps the pair readable.
:::
::::

::: {.takeaway-strip}
Use `::: {.layout-grid-2}` for a comparison or prompt pair—not merely to decorate an ordinary list.
:::

## Layout: vertical stack

:::: {.layout-stack}
::: {.box-question}
First related teaching element.
:::

::: {.box-result}
Second element with consistent breathing room.
:::

::: {.takeaway-strip}
The default gap is `1.4rem`; override `--layout-gap` on the wrapper only when the rendered composition needs it.
:::
::::

## Layout: centred stack

:::: {.layout-stack .layout-center-y}
::: {.box-question .text-center}
Which observation would change your conclusion?
:::

::: {.takeaway-strip .text-center}
`.layout-center-y` centres the short stack in normal flow without transforming its children.
:::
::::

